---
title: "Class8"
format: pdf
author: Mahsa Naeimi
---

# MINI PROJECT

First, we download and import the data to R studio:

```{r}
url <- "https://marcos-diazg.github.io/BIMM143_SP23/class-material/class8/WisconsinCancer.csv"
wisc.df <- read.csv(url, row.names=1)

```

We need to save the input to the project directory and we store the data as `wisc.df`

```{r}
head(wisc.df)
```

we don't want to use the first column, so we need to exclude it:

```{r}
wisc.data <- wisc.df[,-1]
```

Now, form a new data frame and call it `diagnosis`:

```{r}
diagnosis <- as.factor(wisc.df$diagnosis)
```

**Q1**. How many observations are in this dataset?

```{r}
nrow(wisc.data)
```

the number of observations are 569

Q2. How many of the observations have a malignant diagnosis?

```{r}
table(diagnosis)
```

212 of the observations are malignant.

Now, we need to create a new vector for `diagnosis` column

Q3: How many variables/features in the data are suffixed with `_mean`?

```{r}
 grep("_mean", colnames(wisc.df))
```

# Principal Component Analysis

to perform PCA, we need to use `colMean` and `apply` :

```{r}
colMeans(wisc.data)
```

```{r}
apply(wisc.data,2,sd)
```

then, we complete the PCA with `prcomp` function :

```{r}
wisc.pr <- prcomp(scale(wisc.data))
wisc.pr
```

here is the summary of the result by using `summary` function :

```{r}
summary(wisc.pr)
```

**Q4**. From your results, what proportion of the original variance is captured by the first principal components (PC1)?

**Q5**. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?

**Q6**. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?

## Interpreting PCA results

Now we can generate some visualization:

```{r}
biplot(wisc.pr)
```

Q7. What stands out to you about this plot? Is it easy or difficult to understand? Why?

the plot is not easy to interpret because it's very messy and the datapoints are not clear.

Let's generate more standard scatter plot for PC1 and PC2:

```{r}
plot( wisc.pr$x[,1:2] , col = diagnosis , 
     xlab = "PC1", ylab = "PC2")
```

Q8. Generate a similar plot for principal components 1 and 3. What do you notice about these plots?

```{r}
plot(wisc.pr$x[,c(1,3)], col = diagnosis, 
     xlab = "PC1", ylab = "PC3")

```

```{r}
View (wisc.pr$x)
```
