---
title: "Class11HW"
author: "Mahsa Naeimi"
format: pdf
---

Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes.

```{r}
# First we need to read the data from url:
url <- "https://bioboot.github.io/bimm143_S18/class-material/rs8067378_ENSG00000172057.6.txt"
genot <- read.table(url)
```

```{r}
# to find the sample size we use the summary function to get additional stats for each column
summary(genot)
```

There are 462 sample.

```{r}
table(genot$geno)['G/G']
```

From the data table we can see there are 3 genotype: G/G , A/A , G/A

```{r}
# Sample size and Median expression level for G/G genotype
table(genot$geno)['G/G']

summary( genot$exp[ genot$geno == "G/G" ] )
```

```{r}
# Sample size and Median expression level for A/A genotype
table(genot$geno)['A/A']

summary( genot$exp[ genot$geno == "A/A" ] )
```

```{r}
# Sample size and Median expression level for A/G genotype
table(genot$geno)['A/G']

summary( genot$exp[ genot$geno == "A/G" ] )
```

Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3?

```{r}
# Boxplot for each genotype expression data

p <- boxplot(exp~geno, data=genot, xlab="rs8067378 genotype", ylab="ENSG00000172057.4 (RPKM)", notch=T)
```

```{r}
p
```

The boxplot shows that relative expression value of A/A is more than G/G. For this sample, this means higher expression level of ORMDL3 can happen when the genotype is A/A. So, SNP impacts the expression level of ORMDL3. someone with A/A genotype have higher expression level of ORMDL3 than someone with G/G genotype.

```{r}
# to get a better visualization of expression level in each genotype we can use ggplot

library(ggplot2)
ggplot(genot, aes(geno, exp)) + geom_boxplot()
ggplot(genot, aes(geno, exp, fill=geno)) + 
  geom_boxplot(notch=TRUE, outlier.shape = NA) + 
  geom_jitter(shape=16, position=position_jitter(0.2), alpha=0.4)
```
