---
title: "Class 04"
author: "Mahsa Naeimi"
date: 2023-04-19
format: html
editor: 
  markdown: 
    wrap: 72
---

Base R plotting

We are going to start by generating the plot of class 04. This code is
plotting the cars dataset.

```{r}
plot(cars)
```

Ggplot2

First, we need to install the package. We do this by using the
install.packages command.

```{r}
# install.packages('ggplot2')
```

After that, we need to load the package.

```{r}
library(ggplot2)
```

We are going to build the plot of the cars dataframe by using ggplot2.

```{r}
ggplot(data=cars)
ggplot(data = cars) + aes(x=speed, y=dist) 
ggplot(data = cars) + aes(x=speed, y=dist) +geom_point() 
ggplot(data = cars) + aes(x=speed, y=dist) +geom_point()+geom_smooth(method = 'lm')+ labs(title = 'distance vs. speed', subtitle = 'A nice subtitle', caption = 'BIMM143', x= 'Speed (MPH)', y= 'Distance (feet)')
```

Q1

For which phases is data visualization important in our scientific
workflows?

For exploratory Data Analysis,

Plotting Gene Expression data

Loading the data from the URL.

```{r}
 url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt" 
genes <- read.delim(url) 
head(genes)
```

Initial ggplot

```{r}
ggplot(data = genes) + aes(x=Condition1, y=Condition2) + geom_point()


```

Q3

```{r}
nrow(genes) 
ncol(genes) 
colnames(genes)
table(genes[,'State'])
```

Adding color to the plot

```{r}
 p1 <- ggplot(data = genes) + aes(x=Condition1, y=Condition2, col=State) + geom_point() 
 p1
```

Let's change the color scheme.

```{r}
p2 <-p1+ scale_color_manual(values = c('blue', 'grey', 'red')) 
p2
```

Let's add some labels

```{r}
p2+ labs(title = 'Differential gene expression', x='control (no drug)', y= 'Drug treatment', caption= 'BIMM 143 - Class 05')
```
