---
title: "HomeWork6"
author: "Mahsa Naeimi"
date: "2023-05-24"
output:
  pdf_document: default
  html_document: default
---

### **Section 1: Improving analysis code by writing functions**

#### B.

```{r}
# Can you improve this analysis code?
library(bio3d)
# s1 <- read.pdb("4AKE")  # kinase with drug
# s2 <- read.pdb("1AKE")  # kinase no drug
# s3 <- read.pdb("1E4Y")  # kinase with drug
# s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
# s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
# s3.chainA <- trim.pdb(s1, chain="A", elety="CA")
# s1.b <- s1.chainA$atom$b
# s2.b <- s2.chainA$atom$b
# s3.b <- s3.chainA$atom$b
# plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
# plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
# plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")

#improving copy paste issues:
s1 <- read.pdb("4AKE")  # kinase with drug
s2 <- read.pdb("1AKE")  # kinase no drug
s3 <- read.pdb("1E4Y")  # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse = s1.chainA, typ = "l", ylab = "Bfactor")
plotb3(s2.b, sse = s2.chainA, typ = "l", ylab = "Bfactor")
plotb3(s3.b, sse = s3.chainA, typ = "l", ylab = "Bfactor")


```

We want to create a function to be able to visualize protein-drug interactions from PDB data:

```{r}
pr_dr <- function(file, chain, elmnt, fctr) {
   plot_colors <- c("red", "blue", "green")
     for (i in 1:length(file)) {
  s1 <- read.pdb(file[i])

  s1.chain <- trim.pdb(s1, chain = chain, elety = elmnt)
  
  atom_df <- s1.chain$atom
   s1.fctr <- atom_df[, fctr] 
     if (i == 1) {
    plotb3(s1.fctr, sse = s1.chain, typ = "l", ylab = paste(toupper(fctr), "factor", sep = ""), col = plot_colors[i])
     } else {
    lines(s1.fctr, col = plot_colors[i])
  }
     }
     legend("topright", title = "PDB File Name", file, fill = plot_colors, horiz=TRUE, cex = 0.5, inset = c(0.03, 0.06))
}
```

```{r}
# Testing the function for theses 3 protein: 
files <- c("4AKE", "1AKE", "1E4Y")
chains <- "A"
elements <- "CA"
factors <- "b"

pr_dr(files, chains, elements, factors)
```

**Q1. What type of object is returned from the read.pdb() function?** This function returns information about the structure of the protein, its building block, and sequence of the protein from PDB

```{r}
read.pdb("4AKE") 

```

**Q2. What does the trim.pdb() function do?** This function trims the pdb object by selecting specific chains, residues, or atoms based on the provided parameters. In this case, **`trim.pdb()`** is used to extract a specific chain ("A") and element ("CA") from the pdb object.

```{r}
trim.pdb(s1, chain="A", elety="CA")
```

**Q3. What input parameter would turn off the marginal black and grey rectangles in the plots and what do they represent in this case?** To remove the black and grey rectangles, we need delete SSE( secondary structral element) from the chunk.

```{r}
plotb3(s1.b, typ = "l", ylab = "Bfactor")
```

**Q4. What would be a better plot to compare across the different proteins?** a single plot with different colored lines representing each protein's B-factor trends which allows us to compare multiple proteins at the same time.

**Q5. Which proteins are more similar to each other in their B-factor trends. How could you quantify this?** From the clustering we can see that 1E4Y and 1AKE are the most similar ones.

```{r}
bfactors <- rbind(s1.b, s2.b, s3.b)
dist_matrix <- dist(bfactors)
hclust_result <- hclust(dist_matrix)
plot(hclust_result)
```

**Q6. How would you generalize the original code above to work with any set of input protein structures?**

```{r}
pr_dr <- function(file, chain, elmnt, fctr) {
   plot_colors <- c("red", "blue", "green")
     for (i in 1:length(file)) {
  s1 <- read.pdb(file[i])

  s1.chain <- trim.pdb(s1, chain = chain, elety = elmnt)
  
  atom_df <- s1.chain$atom
   s1.fctr <- atom_df[, fctr] 
     if (i == 1) {
    plotb3(s1.fctr, sse = s1.chain, typ = "l", ylab = paste(toupper(fctr), "factor", sep = ""), col = plot_colors[i])
     } else {
    lines(s1.fctr, col = plot_colors[i])
  }
     }
     legend("topright", title = "PDB File Name", file, fill = plot_colors, horiz=TRUE, cex = 0.5, inset = c(0.03, 0.06))
}
```

		

	
